---
title: "GAM_LSA"
author: "Dewan Mohammad Enamul Haque Haque & Ritu Roy"
date: "2024-10-13"
output: html_document
---

## INITIAL SETTINGS 
# Install necessary libraries & load packages
```{r}

list.packages = c("sf", "tidyverse", "pROC", "mapview", "mgcv", "sperrorest", "paletteer")
new.packages = list.packages[!(list.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
```

# loading packages
# this command will load the previously install packages
```{r}
lapply(list.packages, require, character.only=T)
remove(list.packages, new.packages)

```

## Setting up directory

```{r}
setwd("C:/Users/Dewan/Desktop/Testrun/Testrun")
```

# LOADING DATA 
```{r}
Inventory = sf::st_read("C:/Users/Dewan/Desktop/Testrun/Testrun/Inventory2018_2020combined.gpkg")
Inventory2018 = sf::st_read("C:/Users/Dewan/Desktop/Testrun/Testrun/Inventory2018.gpkg")
Boundary = sf::st_read("C:/Users/Dewan/Desktop/Testrun/Testrun/campboundary46n.gpkg")
```
# load mapping units
```{r}
SU = sf::st_read("C:/Users/Dewan/Desktop/Testrun/Testrun/Slopeunitstheend.gpkg")
```
# visualize data
# Save as webpage
# mapview will allow to visualize the available data in an interactive interface.
# the basemap can be switched to topographic maps and satellite imagery 
# stable slope units without landslide are represented with "0"
# unstable slope units with landslides are represented with "1"
```{r}
mapview(SU, zcol="Landslid_2", col.regions=c("dodgerblue1", "firebrick1")) +
  mapview(SU, zcol="Landslid_1", col.regions=c("dodgerblue4", "firebrick1")) 

mapview(SU, zcol="Landslide2", col.regions=c("dodgerblue1", "firebrick1")) +
mapview(SU, zcol="LandslideC", col.regions=c("dodgerblue1", "firebrick1"))
```
## Exploratory data analysis
# histograms and boxplots
```{r}
hist(SU$SlopeM, breaks = 100, xlab="", ylab="Frequency", main = "Average slope (°)")
boxplot(SU$SlopeM ~ SU$Landslid_2, col = c("dodgerblue1", "firebrick1"), main = "Average slope (°)", xlab="", ylab="")
boxplot(SU$AspectM ~ SU$Landslid_2, col = c("dodgerblue1", "firebrick1"), main = "Average Aspect (°)", xlab="", ylab="")
boxplot(SU$RelRelfM ~ SU$Landslid_2, col = c("dodgerblue1", "firebrick1"), main = "Relative relief", xlab="", ylab="")
boxplot(SU$Rain2013_5 ~ SU$Landslid_2, col = c("dodgerblue1", "firebrick1"), main = "Average Rainfall (mm)", xlab="", ylab="")
```
# we can also check patterns using the Probability Density function for both stable and unstable catchments
```{r}
plot(density(SU$SlopeM[SU$Landslid_1==2]), col = "blue", "Average slope (°)", lwd = 2)
plot(density(SU$SlopeM[SU$Landslid_1==1]), col = "blue", "Average slope (°)", lwd = 2)
```
## MODELING ----------------------------------------------------------------
# Statistical modeling using 
# Generalized Additive Modes (GAMs) in the package mgcv https://cran.r-project.org/web/packages/mgcv/index.html
# Frequentist framework

# before we start let us have a look at the total of events in each inventory
```{r}
nrow(Inventory)
```
# model fit
# formula
# bs controls the smooth types to be use in the model and "tp" is used by default.
# k-1 or (k) sets the upper limit on the degrees of freedom associated with a smooth. 
# for more details go to https://cran.r-project.org/web/packages/mgcv/mgcv.pdf or
# (Wood, 2017) https://www.taylorfrancis.com/books/mono/10.1201/9781315370279/generalized-additive-models-simon-wood

#//////////////////////////////////////
## Model for 2021(Monsoon) and prior
#//////////////////////////////////////
```{r}
formula_5000 = Landslid_2 ~
  s(SlopeM, bs="tp", k = 26) + #s() is used to declare the non-linear properties
  s(SlopeR, bs="tp", k = 26)+
  s(SlopeSTD, bs="tp", k = 26)+
  s(SoildepthR, k = 26)+
  s(SoildepthM, k = 26)+
  s(SoilDepthS, k = 26)+
  s(AspectM, k = 26)+
  s(AspectSTD, k = 26)+
  s(PlanCurvR, k = 26)+
  s(PlanCurvM, k = 26)+
  s(ProfCurvR, k = 26)+
  s(ProfCurvM,k = 26)+
  s(RelRelfR, k = 26)+
  s(RelRelfM, k = 26)+
  s(TWIM, k = 26)+
  s(TWIR, k = 26)+
  s(TWISTD, k = 26)+
  s(DistRR, k = 26)+
  s(DistRM, k = 26)+
  s(Rain201321, k = 26)+
  s(Rain2013_4, k = 26)+
  s(SPIR, k = 26)+
  s(SPIM, k = 26)+
  s(SPISTD, k = 26)+
  LULC201720+
  NDVI201720+
  SoilType
```
# Fit the model
```{r}
start.time = Sys.time()
mod_5000 = mgcv::gam(formula_5000, family = binomial, method="REML", data = SU)
end.time = Sys.time()
print(end.time-start.time)
```
# summary will give an overview of the fit
```{r}
summary(mod_5000)
library(Metrics)
bias(mod_5000$fitted.values, (SU$Landslid_2))
ae(mod_5000$fitted.values, (SU$Landslid_2))
mae(mod_5000$fitted.values, (SU$Landslid_2))
rmse(mod_5000$fitted.values, (SU$Landslid_2))
cor(mod_5000$fitted.values, (SU$Landslid_2), method = "pearson")

```

## partial effects
# Visualization 3
# Draw the plot for the specific term and add theme
```{r}
plotss <- draw(mod_5000, select = "s(SlopeSTD)") +
  ylim(c(-10, 10)) +
  theme(axis.text = element_text(size = 25))
plotss
```
# Fitting performance:goodness of fit
# to create a new column in your dataset where the probability is stored
```{r}
library(pROC)
SU$probability = as.numeric(predict(mod_5000, type="response", newdata=SU)) 
myroc_5000 = roc(response=SU$Landslid_2, predictor=SU$probability, auc=T) 
par(pty="s")
plot(myroc_5000, main = round(myroc_5000$auc, 5))
print(myroc_5000$auc)
```
## validation 
# 10-fold cross-validation
# we divide the dataset into 10 equal parts. We will use 9 of those parts to fit our model and validate over the remaining one
# in a interactive way
# partitions
# the library only works with centroids, but this step is merely for visualization
# what we only need is to defined the partitions
# Loading relevant packages
```{r}
library(dplyr)
library(tidyverse)

centroids = dplyr::select(SU, CoordX_1, CoordY_1) %>% sf::st_drop_geometry()
partition = partition_cv(SU, nfold = 10, repetition = 1, seed1 = 123) 
plot(partition, centroids, coords = c("CoordX_1", "CoordY_1"), cex = 0.01, pch = 19)
```
# settings for the loop
# this loop will fit the model over 9 of the defined folds and will predict over the remaining fold
```{r}
fold = (1:10)
SU$prediction = NA
SU_myroc = c(NA)
df = SU
roc = list()
```

# loop
```{r}
for (i in fold){
  id.holdout = partition[[1]][[i]]$test
  df_train = SU[-id.holdout,]
  df_test = SU[id.holdout, ]
  fit = mgcv::gam(formula_5000, data=df_train, family=binomial, method="REML")
  SU$prediction[id.holdout] = as.numeric(predict(fit, type="response", newdata=df_test))
  roc[[i]] = roc(response=df_test$Landslid_2, predictor=SU$prediction[id.holdout], auc=T)
  SU_myroc[i] = as.numeric(unlist(roc[[i]][9]))
}
```
# Let's check the predictive performance over the 10 testing folds
```{r}
boxplot(SU_myroc)
round(SU_myroc,5)
mean(SU_myroc)
median(SU_myroc)
```
# we can plot the 10 roc curves, one generated for every testing fold
```{r}
par(pty="s")
plot(roc[[1]]); plot(roc[[2]], add=T); plot(roc[[3]], add=T); plot(roc[[4]], add=T); plot(roc[[5]], add=T)
plot(roc[[6]], add=T); plot(roc[[7]], add=T); plot(roc[[8]], add=T); plot(roc[[9]], add=T); plot(roc[[10]], add=T)
```
## VISUALIZATION 
# to visualize the results of the fit and the predictions
# we can adjust the strect and type of palette according to your interests
```{r}
mapview(SU, zcol="probability", col.regions=paletteer::paletteer_d("RColorBrewer::RdYlGn", direction=-1))+
  mapview(SU, zcol="prediction", col.regions=paletteer::paletteer_d("RColorBrewer::RdYlGn", direction=-1))+
  mapview(SU, zcol="Landslid_1", col.regions=c("dodgerblue1", "firebrick1"))

mapview(SU, zcol="probability", col.regions=paletteer::paletteer_d("RColorBrewer::RdYlGn", direction=-1), at=seq(0, 1, by=0.2), layer.name="Landslide Susceptibility 2021")
```
##Export Probability & Prediction
```{r}
write.csv(SU$probability, file="probability2021.csv", row.names= FALSE)

write.csv(SU$prediction, file="prediction2020.csv", row.names= FALSE)
```
#//////////////////////////////////////
## Model for 2018 (Monsoon) and prior
#///////////////////////////////////////

```{r}
formula_5000_a = Landslide2 ~
  s(SlopeM, bs="tp", k = 24) + #s() is used to declare the non-linear properties
  s(SlopeR, bs="tp", k = 24)+
  s(SlopeSTD, bs="tp", k = 24)+
  s(SoildepthR, k = 24)+
  s(SoildepthM, k = 24)+
  s(SoilDepthS, k = 24)+
  s(AspectM, k = 24)+
  s(AspectSTD, k = 24)+
  s(PlanCurvR, k = 24)+
  s(PlanCurvM, k = 24)+
  s(ProfCurvR, k = 24)+
  s(ProfCurvM,k = 24)+
  s(RelRelfR, k = 24)+
  s(RelRelfM, k = 24)+
  s(TWIM, k = 24)+
  s(TWIR, k = 24)+
  s(TWISTD, k = 24)+
  s(Rain201317, k = 24)+
  s(Rain2013_1, k = 24)+
  s(SPIR, k = 24)+
  s(SPIM, k = 24)+
  s(SPISTD, k = 24)+
  LULC2017_1+
  NDVI2017_1+
  SoilType
  
```
# Fit the Model
```{r}
mod_5000_a = mgcv::gam(formula_5000_a, family = binomial, method="REML", data = SU)
```
# Model Summary
```{r}
summary(mod_5000_a) 
gam.check(mod_5000_a)
```
# Goodness of fit
```{r}
library(Metrics)
bias(mod_5000_a$fitted.values, (SU$Landslid_2))
ae(mod_5000_a$fitted.values, (SU$Landslid_2))
#plot(density(ae(Fit.size$fitted.values, log(data4fit$ExpandedSlide))))
mae(mod_5000_a$fitted.values, (SU$Landslid_2))
rmse(mod_5000_a$fitted.values, (SU$Landslid_2))
cor(mod_5000_a$fitted.values, (SU$Landslid_2), method = "pearson")

```
## partial effects
# Visualization 3
# Draw the plot for the specific term and add theme

```{r}
plotsm <- draw(mod_5000_a, select = "s(SlopeSTD)") +
  ylim(c(-10, 10)) +
  theme(axis.text = element_text(size = 25))
plotsm
```
# fitting performance
# to create a new column in your dataset where the probability is stored
```{r}
SU$probability_a = as.numeric(predict(mod_5000_a, type="response", newdata=SU)) 
myroc_5000_a = roc(response=SU$Landslide2, predictor=SU$probability_a, auc=T)           
plot(myroc_5000_a, main = round(myroc_5000_a$auc, 5))
print(myroc_5000_a$auc)

```
## validation
# 10-fold cross-validation
# we divide the dataset into 10 equal parts. We will use 9 of those parts to fit our model and validate over the remaining one
# in a interactive way

# partitions
# the library only works with centroids, but this step is merely for visualization
# what we only need is to defined the partitions
```{r}
centroids = dplyr::select(SU, CoordX_1, CoordY_1) %>% sf::st_drop_geometry()
partition = partition_cv(SU, nfold = 10, repetition = 1, seed1 = 123) 
plot(partition, centroids, coords = c("CoordX_1", "CoordY_1"), cex = 0.01, pch = 19)
```
# settings for the loop
# this loop will fit the model over 9 of the defined folds and will predict over the remaining fold
```{r}
fold = (1:10)
SU$prediction_a = NA
SU_myroc_a = c(NA)
df_a = SU
roc_a = list()
```
# loop
```{r}
for (i in fold){
  id.holdout = partition[[1]][[i]]$test
  df_a_train = SU[-id.holdout,]
  df_a_test = SU[id.holdout, ]
  fit_a = mgcv::gam(formula_5000_a, data=df_a_train, family=binomial, method="REML")
  SU$prediction_a[id.holdout] = as.numeric(predict(fit, type="response", newdata=df_a_test))
  roc_a[[i]] = roc(response=df_a_test$Landslide2, predictor=SU$prediction_a[id.holdout], auc=T)
  SU_myroc_a[i] = as.numeric(unlist(roc_a[[i]][9]))
}

```
# let us check the predictive performance over the 10 testing folds
```{r}
boxplot(SU_myroc_a)
round(SU_myroc_a,5)
mean(SU_myroc_a)
median(SU_myroc_a)
```
# we can plot the 10 roc curves, one generated for every testing fold
```{r}
plot(roc_a[[1]]); plot(roc_a[[2]], add=T); plot(roc_a[[3]], add=T); plot(roc_a[[4]], add=T); plot(roc_a[[5]], add=T)
plot(roc_a[[6]], add=T); plot(roc_a[[7]], add=T); plot(roc_a[[8]], add=T); plot(roc_a[[9]], add=T); plot(roc_a[[10]], add=T)
```
# VISUALIZATION 
# to visualize the results of the fit and the predictions
# you can adjust the strect and type of palette according to your interests
```{r}
mapview(SU, zcol="probability_a", col.regions=paletteer::paletteer_d("RColorBrewer::RdYlGn", direction=-1))+
  mapview(SU, zcol="prediction_a", col.regions=paletteer::paletteer_d("RColorBrewer::RdYlGn", direction=-1))+
  mapview(SU, zcol="Landslid_1", col.regions=c("dodgerblue1", "firebrick1"))

mapview(SU, zcol="probability_a", col.regions=paletteer::paletteer_d("RColorBrewer::RdYlGn", direction=-1), at=seq(0, 1, by=0.2), layer.name="Landslide Susceptibility 2018")
```

##Export Probability & Prediction
```{r}
write.csv(SU$probability_a, file="probability2018.csv", row.names= FALSE)

write.csv(SU$prediction_a, file="prediction2018.csv", row.names= FALSE)
```
## Congratulations!!