---
title: "ML_LSA"
author: "Dewan Mohammad Enamul Haque Haque & Ritu Roy"
date: "2024-10-10"
output:
  html_document: default
  pdf_document: default
---

## R Markdown

# R code for landslide susceptibility modeling using Machine Learning Algorithm
# Last modified:  October 2024
# The code is distributed "as is", WITH NO WARRANTY whatsoever!

## Installing/importing required libraries/packages

```{r}
#install.packages("tidyverse")
library(tidyverse)
# Key Libraries
library(lattice)
library(ggplot2)
#install.packages("caret")
library(caret)

```

## Data Import

```{r}
setwd("C:/Users/Dewan/Desktop/Slope Units/R") 
data <-  read.csv("C:/Users/Dewan/Desktop/Slope Units/R/Theend17082023csv.csv", header = T)
```

## Splitting data into training and testing
## Use 60% of data set as training set and remaining 40% as testing set

```{r}
sample <- sample(c(TRUE, FALSE), nrow(data), replace=TRUE, prob=c(0.6,0.4))
train  <- data[sample, ]
test   <- data[!sample, ]

```
## Dimension of data

```{r}
dim(train)
dim(test)

```

## Scale the data for standardization
# Finding the maximum and minimum values except the first column
```{r}
maxs <- apply(train [c(-1)], 2, max)
mins <- apply(train[c(-1)], 2, min)

```

# Scaling the data between 0 and 1
```{r}
scaled_train <- as.data.frame(scale(train[c(-1)], center = mins, scale = maxs - mins))

```
# Combining the target column
```{r}
scaled_train=cbind(scaled_train,train$Landslide)

```
# Renaming the column
```{r}
colnames(scaled_train)[28]<-"Slide2020"

```
##Fit the Model
```{r}
ctrl <- trainControl(method="cv", summaryFunction=twoClassSummary, classProbs=T,savePredictions = TRUE)

```
## Model Training
# Random Forest(rf)
```{r}
tuneGrid.rf <- expand.grid(mtry = c(2, 4, 6, 8))
fit.rf <- train(Slide2020~., data=scaled_train, method="rf", metric="ROC", tuneGrid=tuneGrid.rf, trControl=ctrl)
```
# Support Vector Machine (svmRadial)
```{r}
tuneGrid.svm <- expand.grid(C = c(0.1, 1, 10), sigma = c(0.01, 0.05, 0.1))
fit.svm <- train(Slide2020~., data=scaled_train, method="svmRadial", metric="ROC", tuneGrid=tuneGrid.svm, trControl=ctrl)
```
# Neural Network(nnet)
```{r}
tuneGrid.nnet <- expand.grid(size = c(3, 5, 7), decay = c(0.001, 0.1))
fit.nnet <- train(Slide2020~., data=scaled_train, method="nnet", metric="ROC", tuneGrid=tuneGrid.nnet, trace=FALSE, trControl=ctrl, maxit=100)

```
# XGBoost(xgbLinear)
```{r}
fit.XGB<- train(Slide2020~., data=scaled_train, method="xgbLinear", metric="ROC", trControl=ctrl)

```
# Linear Discriminant Analysis(lda)
```{r}
fit.lda <- train(Slide2020~., data=scaled_train, method="lda", metric="ROC", trControl=ctrl)

```
## Train Model Performance
# Loading relevant packages
```{r}
#install.packages("MLeval")
library(MLeval)

res<-evalm(list(fit.rf),gnames = c("rf"),rlinethick=1,fsize=8)
res<-evalm(list(fit.svm),gnames = c("svm"),rlinethick=1,fsize=8)
res<-evalm(list(fit.lda),gnames = c("lda"),rlinethick=1,fsize=8)
res<-evalm(list(fit.nnet),gnames = c("nnet"),rlinethick=1,fsize=8)
res<-evalm(list(fit.XGB),gnames = c("xgbLinear"),rlinethick=1,fsize=8)
res <- evalm(list(fit.rf, fit.svm, fit.nnet, fit.lda, fit.XGB),
             gnames = c("rf", "svm", "nnet", "lda", "xgbLinear"),
             rlinethick = 1, fsize = 8)

```
# View the optimal results
```{r}
res$optres

```
# Covariate importance
```{r}
varImp(fit.rf)
plot(varImp(fit.rf))
plot(varImp(fit.nnet))
```
## Model Testing
# Sclaing & Renaming
```{r}
maxs <- apply(test [c(-1)], 2, max)
mins <- apply(test[c(-1)], 2, min)
scaled_test <- as.data.frame(scale(test[c(-1)], center = mins, scale = maxs - mins))

scaled_test=cbind(scaled_test,test$Landslide)
colnames(scaled_test)[28]<-"Slide2020"
``` 
# Predictions for the test data
# Loading relevant packages
```{r}
library(ROCR)
library(pROC)

p_rf_test <- predict(fit.rf, scaled_test, type = "prob")
p_svm_test <- predict(fit.svm, scaled_test, type = "prob")
p_nnet_test <- predict(fit.nnet, scaled_test, type = "prob")
p_lda_test <- predict(fit.lda, scaled_test, type = "prob")
p_xgb_test <- predict(fit.XGB, scaled_test, type = "prob")
```
# Plot ROC curves for each model and extract AUC
```{r}
roc_rf <- roc(scaled_test$Slide2020, p_rf_test[,"YES"], plot=TRUE, col="green", lwd=3, legacy.axes=TRUE, main="ROC Curves")
roc_nnet <- roc(scaled_test$Slide2020, p_nnet_test[,"YES"], plot=TRUE, col="blue", lwd=3, legacy.axes=TRUE, add=TRUE)
roc_svm <- roc(scaled_test$Slide2020, p_svm_test[,"YES"], plot=TRUE, col="purple", lwd=3, legacy.axes=TRUE, add=TRUE)
roc_lda <- roc(scaled_test$Slide2020, p_lda_test[,"YES"], plot=TRUE, col="red", lwd=3, legacy.axes=TRUE, add=TRUE)
roc_xgb <- roc(scaled_test$Slide2020, p_xgb_test[,"YES"], plot=TRUE, col="black", lwd=3, legacy.axes=TRUE, add=TRUE)

```
# Dynamically generate AUC values

```{r}
auc_rf <- auc(roc_rf)
auc_nnet <- auc(roc_nnet)
auc_svm <- auc(roc_svm)
auc_lda <- auc(roc_lda)
auc_xgb <- auc(roc_xgb)
```
# Add a legend with dynamically calculated AUC values
```{r}
par(pty="s")
plot(roc_nnet, col="blue", lwd=3, legacy.axes=TRUE, main="ROC Curves")
plot(roc_rf, col="green", lwd=3, add=TRUE)
plot(roc_lda, col="red", lwd=3, add=TRUE)
plot(roc_svm, col="black", lwd=3, add=TRUE)
plot(roc_xgb, col="purple", lwd=3, add=TRUE)


legend("bottomright", legend=c(
  paste("NNET_AUC", round(auc_nnet, 2)),
  paste("RF_AUC", round(auc_rf, 2)),
  paste("LDA_AUC", round(auc_lda, 2)),
  paste("SVM_AUC", round(auc_svm, 2)),
  paste("XGB_AUC", round(auc_xgb, 2))),
  col=c("blue","green","red","black","purple"),
  lty=1, lwd=4, cex=0.8)

```
## Produce LSP map using training model results and raster layers data
# Install necessary libraries & load packages
```{r}
#install.packages("raster")
#install.packages("rgdal")
# Install terra package (replaces rgdal and raster)
#install.packages("terra")

library(terra)
library(raster)
#library(rgdal)

```
# To check file exists
```{r}
file.exists("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/ElevationM.tif")

```
#Load the Raster data
```{r}
ElevationR = raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/ElevationR.tif")
ElevationM = raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/ElevationM.tif")  
SlopeR= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/SlopeR.tif")
SlopeM= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/SlopeM.tif")
SlopeSTD= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/SlopeSTD.tif")
SoildepthR= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/SoildepthR.tif")
SoildepthM= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/SoildepthM.tif")
SoilDepthSTD= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/SoilDepthSTD.tif")
AspectM= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/AspectM.tif")
AspectSTD= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/AspectSTD.tif")
PlanCurvM= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/PlanCurvM.tif")
PlanCurvR= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/PlanCurvR.tif")
ProfCurvM= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/ProfCurvM.tif")
ProfCurvR= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/ProfCurvR.tif")
RelRelfM= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/RelRelfM.tif")
RelRelfR= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/RelRelfR.tif")
TWIM= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/TWIM.tif")
TWIR= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/TWIR.tif")
TWISTD= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/TWISTD.tif")
SoilType= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/SoilType.tif")
DistRM= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/DistRM.tif")
DistRR= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/DistRR.tif")
Rain201321R= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/Rain201321R.tif")
Rain201321MAX= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/Rain201321MAX.tif")
LULC20172021= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/LULC20172021.tif")
NDVI20172021= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/NDVI20172021.tif")
SPIR= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/SPIR.tif")
SPIM= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/SPIM.tif")
SPISTD= raster("C:/Users/Dewan/Desktop/Slope Units/R/MLFactor_test/SPISTD.tif")

```
# check attributes and projection and extent
```{r}
extent(ElevationM)
extent(ElevationR)
extent(SlopeM)
extent(SlopeR)
extent(SlopeSTD)
extent(SoildepthM)
extent(SoildepthR)
extent(AspectM)
extent(AspectSTD)
extent(PlanCurvM)
extent(PlanCurvR)
extent(ProfCurvM)
extent(ProfCurvR)
extent(RelRelfM)
extent(RelRelfR)
extent(TWIM)
extent(TWIR)
extent(TWISTD)
extent(SoilType)
extent(DistRM)
extent(DistRR)
extent(Rain201321MAX)
extent(Rain201321R)
extent(SPIR)
extent(SPIM)
extent(SPISTD)
extent(NDVI20172021)
extent(LULC20172021)
extent(SoilDepthSTD)

```
# If different extent, then resample them using the smallest area
```{r}
DistRR_a <- resample(DistRR,ElevationM, resample='bilinear')
DistRM_a <- resample(DistRM,ElevationM, resample='bilinear')
SoilType_a <- resample(SoilType,ElevationM, resample='bilinear')
TWIM_a <- resample(TWIM,ElevationM, resample='bilinear')
TWIR_a <- resample(TWIR,ElevationM, resample='bilinear')
TWISTD_a <- resample(TWISTD,ElevationM, resample='bilinear')
RelRelfM_a <- resample(RelRelfM,ElevationM, resample='bilinear')
RelRelfR_a <- resample(RelRelfR,ElevationM, resample='bilinear')
NDVI20172021_a <- resample(NDVI20172021,ElevationM, resample='bilinear')
LULC20172021_a <- resample(LULC20172021,ElevationM, resample='bilinear')

```
# check the new extent
```{r}
extent(RelRelfM_a)

```
# write to a new geotiff file
```{r}
writeRaster(DistRR_a,filename="resampled/DistRR.tif", format="GTiff", overwrite=TRUE) 
writeRaster(DistRM_a,filename="resampled/DistRM.tif", format="GTiff", overwrite=TRUE)
writeRaster(SoilType_a,filename="resampled/SoilType.tif", format="GTiff", overwrite=TRUE)
writeRaster(TWIM_a,filename="resampled/TWIM.tif", format="GTiff", overwrite=TRUE)
writeRaster(TWIR_a,filename="resampled/TWIR.tif", format="GTiff", overwrite=TRUE)
writeRaster(TWISTD_a,filename="resampled/TWISTD.tif", format="GTiff", overwrite=TRUE)
writeRaster(RelRelfM_a,filename="resampled/RelRelfM.tif", format="GTiff", overwrite=TRUE)
writeRaster(RelRelfR_a,filename="resampled/RelRelfR.tif", format="GTiff", overwrite=TRUE)
writeRaster(NDVI20172021_a,filename="resampled/NDVI20172021.tif", format="GTiff", overwrite=TRUE)
writeRaster(LULC20172021_a,filename="resampled/LULC20172021.tif", format="GTiff", overwrite=TRUE)

```
## stack multiple raster files
```{r}
Stack_List= list.files(path = "resampled/",pattern = "tif$", full.names = TRUE)

Rasters=stack(Stack_List)
names(Rasters)

```
# scale the raster stack data
```{r}
scaled_raster<-scale(Rasters,center = TRUE,scale = TRUE)

```
# Downscale the resolution of the raster
```{r}
New<-aggregate(scaled_raster,10)

```
## Predict the map
```{r}
nnet<-1-raster::predict(New,fit.nnet,type="prob")
plot(nnet)

```
#save the map
```{r}
proj4string(nnet)=CRS( "+proj=utm +zone=46 +datum=WGS84 +units=m +no_defs")
writeRaster(nnet,filename = "final prediction map/nnet.tif", format="GTiff", overwrite=TRUE)

```
# Congratulations!!

